### 2025-10-27 Text/Thesis/Chapters/Chapter2.qmd
final version thesis

We denote $Z \sim \Gamma_{\mathrm{SAR}}(L, \mu)$ and $Z \sim \mathcal{G}_I^0(\alpha, \gamma, L)$ when the random variable $Z$ follows the respective distributions, characterized by the following probability density functions (pdfs):
\begin{equation}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    = \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z) \label{E:gamma1}
\end{equation}
and
\begin{align}
 f_{\mathcal{G}^0_I}\bigl(z;  \alpha,\gamma, L \bigr) &=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm{1}_{\mathbb{R}_+}(z), \label{E:gi01}
\end{align}
where $\mu > 0$ is the mean intensity, $\gamma > 0$ is a scale parameter, $\alpha < 0$ controls the degree of texture (roughness), $L \geq 1$ is the number of looks (either nominal or estimated, thus not restricted to integer values), $\Gamma(\cdot)$ is
the gamma function, 
and $\mathbbm 1_{A}(z)$ is the indicator function
of the set $A$.

@@

Although the $\mathcal{G}_I^0$ distribution is defined by the parameters $\alpha$ and $\gamma$, in the SAR literature&nbsp;[@Nascimento2010] the texture $\alpha$ and the mean $\mu$ are usually used. Reparameterizing&nbsp;\eqref{E:gi01} with $\mu$, and denoting this model as $Z \sim \mathcal{G}_I^0(\alpha, \mu, L)$ we obtain:
\begin{equation}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    = \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi02}
\end{equation}
The $\Gamma_{\mathrm{SAR}}$ model is a particular case of the $\mathcal{G}^0_I$ distribution, as demonstrated in&nbsp;[@Frery1997]. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) \quad \text{ when } \alpha\to-\infty.
$$

@@

In the context of SAR imaging, homogeneous areas produce fully developed speckle and are effectively modeled by the $Z \sim\Gamma_{\text{SAR}}(L, \mu)$ distribution. In contrast, textured or heterogeneous regions are better represented by the  $Z \sim\mathcal{G}^0_I(\mu, \alpha, L)$ model, where the roughness parameter $\alpha$ controls the degree of heterogeneity. Across this section, we show that all three entropy measures applied to $\mathcal{G}^0_I$ can be decomposed as:
\begin{equation}
\label{eq:decomp-intro}
\boxed{\;
\text{Entropy}\bigl(\mathcal{G}^{0}_{I}\bigr)
\;=\;
\text{Entropy}\bigl(\Gamma_{\mathrm{SAR}}\bigr)
\;+\;
\Delta_{\alpha}
\;}
\end{equation}
where  $\Delta_\alpha$ represents the \textit{excess entropy} induced by the roughness parameter $\alpha$. Notably, this excess vanishes in the homogeneous limit ($\alpha \to -\infty$), reducing \eqref{eq:decomp-intro} to the baseline case.

@@

The parametric representation of Shannon entropy for a system described by a continuous random variable is:
\begin{equation}
  \label{eq:entropy-sh}
  H(Z) =   -\mathbb{E}\bigl[ \ln f(z)\bigr]=-\int_{-\infty }^\infty f(z)\ln f(z)\, \mathrm{d}z,
\end{equation}
where $f(\cdot)$ is the pdf that characterizes the distribution of the real-valued random variable $Z$.

@@

\textbf{Homogeneous case.} For the $\Gamma_{\mathrm{SAR}}$ distribution defined in \eqref{E:gamma1}, we obtain:
\begin{equation}
\label{eq:gamma-sh1}
H\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr) = L - \ln L + \ln\Gamma(L) + (1 - L)\psi^{(0)}(L) + \ln \mu,
\end{equation}
where $\psi^{(0)}(\cdot)$ is the digamma function; $H(\Gamma_{\mathrm{SAR}})$ represents the baseline entropy of fully developed speckle.

\textbf{Textured case.} In the case of the $\mathcal{G}^0_I$ distribution \eqref{E:gi02}, we adapt existing expressions reported by De A. Ferreira and Nascimento&nbsp;[-@Ferreira2020] to our parametrization, using the relation $\mu = -\gamma/(\alpha + 1)$. This yields:
\begin{multline}
\label{eq:GIO-Sh1}
H\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) =H\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr)\;+\;
\Bigl[ (L-\alpha) \psi^{(0)}(L-\alpha)-(1-\alpha)\psi^{(0)}(-\alpha)
+\ln (-1-\alpha) \\
-\ln\Gamma(L-\alpha) 
+\ln\Gamma(-\alpha)-L\Bigr].
\end{multline}
Equation \eqref{eq:GIO-Sh1} can be interpreted as: 
\vspace{-15pt}
$$
H(\mathcal{G}^0_I)
=
\underbrace{H\bigl(\Gamma_{\mathrm{SAR}}\bigr)}_{\text{baseline entropy}}
\hspace{1.8em} + \hspace{-1.0em}
\underbrace{\Delta_\alpha}_{\text{excess entropy caused by texture}}\hspace{-4.0em},
$$
\vspace{-10pt}
where
\vspace{-15pt}
\begin{multline} 
  \Delta_\alpha
  =
    (L-\alpha)\,\psi^{(0)}(L-\alpha)
   -(1-\alpha)\,\psi^{(0)}(-\alpha)
   +\ln(-1-\alpha)
\\[-2pt]\hphantom{\Delta^{\mathrm{Sh}}_\alpha={}}
   -\ln\Gamma(L-\alpha)+\ln\Gamma(-\alpha)-L,
   \label{eq:GIO-Sh2}
\end{multline}
the extra term $\Delta_\alpha$ captures how much entropy increases due to texture (heterogeneity), caused by the roughness parameter $\alpha$. When $\alpha \to -\infty$ (fully developed speckle), the excess term tends to zero, and Equation \eqref{eq:GIO-Sh1} collapses to the homogeneous case in \eqref{eq:gamma-sh1}, i.e.,
\vspace{-15pt}
$$
\lim_{\alpha \to -\infty} H\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr).
$$ 
Appendix&nbsp;\ref{app:A1} provides a proof of this limiting behavior.
<!-- This motivates our work: we check whether the \emph{excess} is statistically different from zero.  -->
In practice, detecting heterogeneity is equivalent to testing whether $\Delta_\alpha$ is significantly different from zero.

### Rényi Entropy
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$ is defined as:
\begin{equation}
\label{E:entropy-R}
R_\lambda(Z)  = \frac{1}{1-\lambda }\ln \Bigl[ \mathbb{E}\Bigl( f^{\lambda - 1}(Z)\Bigr) \Bigr] =\frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \,\mathrm{d}z.
\end{equation}
As $\lambda \to 1$, $R_{\lambda}(Z)$ reduces to the Shannon entropy $H(Z)$.

@@

\textbf{Homogeneous case.} Let $Z\sim\Gamma_{\mathrm{SAR}}(L,\mu)$ with pdf given in \eqref{E:gamma1}. We compute
$$
I = \int_{0}^{\infty}
    \bigl[f_{\Gamma_{\mathrm{SAR}}}(z;L,\mu)\bigr]^{\lambda}
    \,\mathrm{d}z
  = \Bigl(\tfrac{L^{L}}{\Gamma(L)\,\mu^{L}}\Bigr)^{\!\lambda}
    \int_{0}^{\infty} z^{\lambda(L-1)}
    \exp\Bigl(-\tfrac{\lambda L}{\mu}\,z\Bigr)\,\mathrm{d}z.
$$
Using the Gamma integral identity
\begin{align*}
\int_{0}^{\infty} x^{p-1}e^{-q x}\,\mathrm{d}x
&= \frac{\Gamma(p)}{q^{p}}, \quad p,q>0,
\end{align*}
with
\vspace{-10pt}
$$
p = \lambda(L-1)+1,\quad
q = \frac{\lambda L}{\mu},
$$
we obtain
$$
I
= \Bigl(\tfrac{L^{L}}{\Gamma(L)\,\mu^{L}}\Bigr)^{\!\lambda}
  \frac{\Gamma\bigl(\lambda(L-1)+1\bigr)}
       {\bigl(\tfrac{\lambda L}{\mu}\bigr)^{\lambda(L-1)+1}}.
$$
Taking logarithms and simplifying gives
\begin{equation*}
\ln I
= (1-\lambda)\bigl(\ln \mu - \ln L\bigr)
- \lambda\,\ln\Gamma(L)
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)
- \bigl(\lambda(L-1)+1\bigr)\ln\lambda.
\end{equation*}
Substituting into \eqref{E:entropy-R} yields the closed-form
\begin{equation}
  \label{eq-GammaSAR-R}
  \boxed{
  R_{\lambda}\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)
  = \ln\mu - \ln L
+ \frac{1}{1-\lambda}
\Bigl[
-\lambda\,\ln\Gamma(L)
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)
- \bigl(\lambda(L-1)+1\bigr)\ln\lambda
\Bigr].
}
\end{equation}

\textbf{Textured case.} Starting from the definition of Rényi entropy in  \eqref{E:entropy-R} and the pdf of  $Z\sim\mathcal{G}^0_I(\alpha,\gamma,L)$ given in \eqref{E:gi01}, we define:
$$
I
= \int_{0}^{\infty}\bigl[f_{\mathcal{G}^0_I}(z;\alpha,\gamma,L)\bigr]^{\lambda}\,\mathrm{d}z
= C^{\lambda}
  \int_{0}^{\infty}
    \frac{z^{\lambda(L-1)}}{(\gamma+Lz)^{\,\lambda(L-\alpha)}}\,\mathrm{d}z,
\qquad
C=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}.
$$
The parameterisation satisfies $\gamma=-\mu(\alpha+1)$, so the final result will be expressed in terms of $\mu$.

@@

This expression extends the Rényi entropy of the $\Gamma_{\mathrm{SAR}}(L,\mu)$ model \eqref{eq-GammaSAR-R} by including additional terms that depend on the texture parameter $\alpha$. Specifically:

@@

&=R_{\lambda}\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)

@@

@@

\lim_{\alpha \to -\infty} R_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = R_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr).

@@

Appendix&nbsp;\ref{app:A24} provides proof of this limiting behavior.

### Tsallis entropy

For a continuous random variable $Z$ with pdf $f(z)$, the Tsallis entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$ is defined as:
\begin{equation}
\label{eq:tsallis}
T_\lambda(Z) = \frac{1}{\lambda - 1} \mathbb{E} \Bigl[ 1 - (f(Z))^{\lambda - 1} \Bigr]
= \frac{1}{\lambda - 1} \Bigl[ 1 - \int_{-\infty}^{\infty} \bigl(f(z)\bigr)^{\lambda} \,\mathrm{d}z \Bigr]. 
\end{equation}
In the limit $\lambda\to1$,
  $T_\lambda(Z)\to H(Z)$, making Tsallis a
  one-parameter generalization of Shannon entropy.

@@

\textbf{Homogeneous case.} 
Let $Z\sim\Gamma_{\mathrm{SAR}}(L,\mu)$  with pdf given by \eqref{E:gamma1}. Define the integral:
\begin{equation*}
I
= \int_0^\infty \left[f_{\Gamma_{\text{SAR}}}(z;L,\mu)\right]^{\lambda} \,\mathrm{d}z.
\end{equation*}
Substituting the density function and simplifying constants, we get
\begin{equation*}
I
= \left(\frac{L^L}{\Gamma(L)\,\mu^L}\right)^{\lambda}
  \int_0^\infty
     z^{\lambda(L-1)}
     \exp\left(-\frac{\lambda L}{\mu} z\right)\,\mathrm{d}z.
\end{equation*}
Using the Gamma integral identity
\begin{align*}
\int_{0}^{\infty} x^{p-1}e^{-q x}\,\mathrm{d}x
&= \frac{\Gamma(p)}{q^{p}}, \quad p,q>0,
\end{align*}
with $p = \lambda(L-1) + 1$ and $q = \frac{\lambda L}{\mu}$. This leads to
\begin{equation*}
I
= \left(\frac{L^L}{\Gamma(L)\,\mu^L}\right)^{\lambda}
  \cdot
  \frac{\Gamma\big(\lambda(L-1)+1\big)}
       {\left(\frac{\lambda L}{\mu}\right)^{\lambda(L-1)+1}}.
\end{equation*}
After rearranging powers of $L$, $\mu$ and $\lambda$, we obtain
\begin{equation}
I
= \frac{L^{\lambda - 1}\,\mu^{1 - \lambda}\;
       \Gamma\big(\lambda(L - 1) + 1\big)}
      {\lambda^{\lambda(L - 1) + 1}\,
       \left[\Gamma(L)\right]^{\lambda}}.
\label{eq:Jgamma1}
\end{equation}
Substituting Equation \eqref{eq:Jgamma1} into the Tsallis entropy definition \eqref{eq:tsallis}, we have:
\begin{equation}
T_\lambda\big(\Gamma_{\text{SAR}}(L,\mu)\big)
= \frac{1}{\lambda - 1}
   \left[
     1 -
     \frac{L^{\lambda - 1}\,\mu^{1 - \lambda}\;
           \Gamma\big(\lambda(L - 1) + 1\big)}
          {\lambda^{\lambda(L - 1) + 1}\;
           \left[\Gamma(L)\right]^{\lambda}}
   \right].
\label{eq:TsallisGammaFinal1}
\end{equation}
This expression can also be written in an equivalent form:
<!-- \begin{equation*} -->
<!-- T_\lambda\big(\Gamma_{\text{SAR}}(L,\mu)\big) -->
<!-- = \frac{1}{\lambda - 1} \left[ -->
<!--    1 - -->
<!--    \mu^{1 - \lambda}\, -->
<!--    L^{\lambda - 1}\, -->
<!--    \lambda^{ - [\lambda(L - 1) + 1] }\, -->
<!--    \Gamma\big(\lambda(L - 1) + 1\big)\, -->
<!--    \Gamma(L)^{ - \lambda} -->
<!-- \right], -->
<!-- \end{equation*} -->
\begin{equation}
\label{eq:gammasar-Tsallis}
\boxed{%
\begin{aligned}
T_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)=
\frac{1}{\lambda-1}\Bigl\{1-
\exp\Bigl[
(1-\lambda)\ln\mu
+(\lambda-1)\ln L
+\ln\Gamma\bigl(\lambda(L-1)+1\bigr) \\
-\lambda\ln\Gamma(L)
-(\lambda(L-1)+1)\ln\lambda
\Bigr]\Bigr\}.
\end{aligned}}
\end{equation}

@@

To express $J$ in terms of the mean $\mu$, we use the relation $\gamma = -\mu(\alpha + 1)$ valid for $\alpha < -1$. Substituting into \eqref{eq:J_beta1} gives

@@

This expression extends the Tsallis entropy of the $\Gamma_{\mathrm{SAR}}(L,\mu)$ model \eqref{eq:gammasar-Tsallis} by including additional terms that depend on the texture parameter $\alpha$. It can be written in an exponential-logarithmic form as:

@@

  T_\lambda\bigl(\mathcal{G}^0_I(\mu,\alpha,L)\bigr) = T_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)\;+\;\frac{1}{\lambda-1}\,

@@

\lim_{\alpha \to -\infty} T_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = T_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr).

@@

Appendix&nbsp;\ref{app:A33} provides proof of this limiting behavior.

>>>>

We denote $Z \sim \Gamma_{\mathrm{SAR}}(L, \mu)$ and $Z \sim \mathcal{G}_I^0(\alpha, \gamma, L)$ when the random variable $Z$ follows the respective distributions, characterized by the following probability density functions (pdfs):
\begin{equation}
	f_{\Gamma_{\mathrm{SAR}}}\bigl(z;\mu, L\bigr) 
    = \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z) \label{E:gamma1}
\end{equation}
and
\begin{align}
 f_{\mathcal{G}^0_I}\bigl(z;  \alpha,\gamma, L \bigr) &=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm{1}_{\mathbb{R}_+}(z), \label{E:gi01}
\end{align}
where $\mu > 0$ is the mean intensity, $\gamma > 0$ is a scale parameter, $\alpha < 0$ controls the degree of texture (roughness), $L \geq 1$ is the number of looks (either nominal or estimated, thus not restricted to integer values), $\Gamma(\cdot)$ is
the gamma function, 
and $\mathbbm 1_{A}(z)$ is the indicator function
of the set $A$.

@@

Although the $\mathcal{G}_I^0$ distribution is defined by the parameters $\alpha$ and $\gamma$, in the SAR literature&nbsp;[@Nascimento2010] the texture $\alpha$ and the mean $\mu$ are usually used. Reparameterizing&nbsp;\eqref{E:gi01} with $\mu$, and denoting this model as $Z \sim \mathcal{G}_I^0(\alpha, \mu, L)$ we obtain:
\begin{equation}
    f_{\mathcal{G}^0_I}\bigl(z;  \alpha, \mu, L \bigr) 
    = \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi02}
\end{equation}
The $\Gamma_{\mathrm{SAR}}$ model is a particular case of the $\mathcal{G}^0_I$ distribution, as demonstrated in&nbsp;[@Frery1997]. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z;  \alpha, \mu, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z; \mu, L) \quad \text{ when } \alpha\to-\infty.
$$

@@

In the context of SAR imaging, homogeneous areas produce fully developed speckle and are effectively modeled by the $Z \sim\Gamma_{\text{SAR}}(L, \mu)$ distribution. In contrast, textured or heterogeneous regions are better represented by the  $Z \sim\mathcal{G}^0_I( \alpha, \mu, L)$ model, where the roughness parameter $\alpha$ controls the degree of heterogeneity. Across this section, we show that all three entropy measures applied to $\mathcal{G}^0_I$ can be decomposed as:
\begin{equation}
\label{eq:decomp-intro}
\boxed{\;
\text{Entropy}\bigl(\mathcal{G}^{0}_{I}\bigr)
\;=\;
\text{Entropy}\bigl(\Gamma_{\mathrm{SAR}}\bigr)
\;+\;
\Delta_{\alpha}
\;}
\end{equation}
where  $\Delta_\alpha$ represents the \textit{excess entropy} induced by the roughness parameter $\alpha$. Notably, this excess vanishes in the homogeneous limit ($\alpha \to -\infty$), reducing \eqref{eq:decomp-intro} to the baseline case.

@@

The parametric representation of Shannon entropy for a system described by a continuous random variable is:
\begin{equation}
  \label{eq:entropy-sh}
  H(Z) =   -\mathbb{E}\bigl[ \ln f(z)\bigr]=-\int_{\mathcal{B}} f(z)\ln f(z)\, \mathrm{d}z,
\end{equation}
where $f(\cdot)$ is the pdf that characterizes the distribution of the real-valued random variable $Z$ and $\mathcal{B} \subseteq \mathbb{R}$ denotes the support of $f(z)$.

@@

\textbf{Homogeneous case.} For the $\Gamma_{\mathrm{SAR}}$ distribution defined in \eqref{E:gamma1}, we obtain:
\begin{equation}
\label{eq:gamma-sh1}
H\bigl(\Gamma_{\mathrm{SAR}}(\mu, L)\bigr) = L - \ln L + \ln\Gamma(L) + (1 - L)\psi^{(0)}(L) + \ln \mu,
\end{equation}
where $\psi^{(0)}(\cdot)$ is the digamma function; $H(\Gamma_{\mathrm{SAR}})$ represents the baseline entropy of fully developed speckle.

\textbf{Textured case.} In the case of the $\mathcal{G}^0_I$ distribution \eqref{E:gi02}, we adapt existing expressions reported by De A. Ferreira and Nascimento&nbsp;[-@Ferreira2020] to our parametrization, using the relation $\mu = -\gamma/(\alpha + 1)$. This yields:
\begin{multline}
\label{eq:GIO-Sh1}
H\bigl(\mathcal{G}^0_I(\alpha, \mu, L)\bigr) =H\bigl(\Gamma_{\mathrm{SAR}}(\mu, L)\bigr)\;+\;
\Bigl[ (L-\alpha) \psi^{(0)}(L-\alpha)-(1-\alpha)\psi^{(0)}(-\alpha)
+\ln (-1-\alpha) \\
-\ln\Gamma(L-\alpha) 
+\ln\Gamma(-\alpha)-L\Bigr].
\end{multline}
Equation \eqref{eq:GIO-Sh1} can be interpreted as: 
\vspace{-15pt}
$$
H(\mathcal{G}^0_I)
=
\underbrace{H\bigl(\Gamma_{\mathrm{SAR}}\bigr)}_{\text{baseline entropy}}
\hspace{1.8em} + \hspace{-1.0em}
\underbrace{\Delta_\alpha}_{\text{excess entropy caused by texture}}\hspace{-4.0em},
$$
\vspace{-10pt}
where
\vspace{-15pt}
\begin{multline} 
  \Delta_\alpha
  =
    (L-\alpha)\,\psi^{(0)}(L-\alpha)
   -(1-\alpha)\,\psi^{(0)}(-\alpha)
   +\ln(-1-\alpha)
\\[-2pt]\hphantom{\Delta^{\mathrm{Sh}}_\alpha={}}
   -\ln\Gamma(L-\alpha)+\ln\Gamma(-\alpha)-L,
   \label{eq:GIO-Sh2}
\end{multline}
the extra term $\Delta_\alpha$ captures how much entropy increases due to texture (heterogeneity), caused by the roughness parameter $\alpha$. When $\alpha \to -\infty$ (fully developed speckle), the excess term tends to zero, and Equation \eqref{eq:GIO-Sh1} collapses to the homogeneous case in \eqref{eq:gamma-sh1}, i.e.,
\vspace{-15pt}
$$
\lim_{\alpha \to -\infty} H\bigl(\mathcal{G}^0_I( \alpha, \mu, L)\bigr) = H\bigl(\Gamma_{\mathrm{SAR}}(\mu, L)\bigr).
$$ 
Appendix A provides a proof of this limiting behavior.
<!-- This motivates our work: we check whether the \emph{excess} is statistically different from zero.  -->
In practice, detecting heterogeneity is equivalent to testing whether $\Delta_\alpha$ is significantly different from zero.

### Rényi Entropy
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$ is defined as:
\begin{equation}
\label{E:entropy-R}
R_\lambda(Z)  = \frac{1}{1-\lambda }\ln \Bigl[ \mathbb{E}\Bigl( f^{\lambda - 1}(Z)\Bigr) \Bigr] =\frac{1}{1 - \lambda} \ln \int_{\mathcal{B}} [f(z)]^\lambda \,\mathrm{d}z,
\end{equation}
where $\mathcal{B} \subseteq \mathbb{R}$ denotes the support of $f(z)$.
As $\lambda \to 1$, $R_{\lambda}(Z)$ reduces to the Shannon entropy $H(Z)$.

@@

\textbf{Homogeneous case.} Let $Z\sim\Gamma_{\mathrm{SAR}}(\mu, L)$ with pdf given in \eqref{E:gamma1}. We compute
I = \int_{0}^{\infty}
    \bigl[f_{\Gamma_{\mathrm{SAR}}}(z;\mu, L)\bigr]^{\lambda}
    \,\mathrm{d}z
  = \Bigl(\tfrac{L^{L}}{\Gamma(L)\,\mu^{L}}\Bigr)^{\!\lambda}
    \int_{0}^{\infty} z^{\lambda(L-1)}
    \exp\Bigl(-\tfrac{\lambda L}{\mu}\,z\Bigr)\,\mathrm{d}z.
Using the Gamma integral identity
\int_{0}^{\infty} x^{p-1}e^{-q x}\,\mathrm{d}x
&= \frac{\Gamma(p)}{q^{p}}, \quad p,q>0,
with
p = \lambda(L-1)+1,\quad
q = \frac{\lambda L}{\mu},
= \Bigl(\tfrac{L^{L}}{\Gamma(L)\,\mu^{L}}\Bigr)^{\!\lambda}
  \frac{\Gamma\bigl(\lambda(L-1)+1\bigr)}
       {\bigl(\tfrac{\lambda L}{\mu}\bigr)^{\lambda(L-1)+1}}.
Taking logarithms and simplifying gives
\ln I
= (1-\lambda)\bigl(\ln \mu - \ln L\bigr)
- \lambda\,\ln\Gamma(L)
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)
- \bigl(\lambda(L-1)+1\bigr)\ln\lambda.
Substituting into \eqref{E:entropy-R} yields the closed-form
\begingroup\small
  \label{eq-GammaSAR-R}
  \boxed{
  R_{\lambda}\bigl(\Gamma_{\mathrm{SAR}}(\mu, L)\bigr)
  = \ln\mu - \ln L
+ \frac{1}{1-\lambda}
-\lambda\,\ln\Gamma(L)
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)
- \bigl(\lambda(L-1)+1\bigr)\ln\lambda
}
\endgroup
\textbf{Textured case.} Starting from the definition of Rényi entropy in  \eqref{E:entropy-R} and the pdf of  $Z\sim\mathcal{G}^0_I(\alpha,\gamma,L)$ given in \eqref{E:gi01}, we define:
= \int_{0}^{\infty}\bigl[f_{\mathcal{G}^0_I}(z;\alpha,\gamma,L)\bigr]^{\lambda}\,\mathrm{d}z
= C^{\lambda}
\qquad
C=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}.
The parameterisation satisfies $\gamma=-\mu(\alpha+1)$, so the final result will be expressed in terms of $\mu$.

@@

This expression extends the Rényi entropy of the $\Gamma_{\mathrm{SAR}}(\mu, L)$ model \eqref{eq-GammaSAR-R} by including additional terms that depend on the texture parameter $\alpha$. Specifically:
\begingroup\small

@@

&=R_{\lambda}\bigl(\Gamma_{\mathrm{SAR}}(\mu, L)\bigr)

@@

\endgroup

@@

\lim_{\alpha \to -\infty} R_\lambda\bigl(\mathcal{G}^0_I( \alpha, \mu, L)\bigr) = R_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr).

@@

Appendix A provides proof of this limiting behavior.

### Tsallis Entropy

For a continuous random variable $Z$ with pdf $f(z)$, the Tsallis entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$ is defined as:
\begin{equation}
\label{eq:tsallis}
T_\lambda(Z) = \frac{1}{\lambda - 1} \mathbb{E} \Bigl[ 1 - (f(Z))^{\lambda - 1} \Bigr]
= \frac{1}{\lambda - 1} \Bigl[ 1 - \int_{\mathcal{B}} \bigl(f(z)\bigr)^{\lambda} \,\mathrm{d}z \Bigr],
\end{equation}
where $\mathcal{B} \subseteq \mathbb{R}$ denotes the support of $f(z)$.
In the limit $\lambda\to1$,
  $T_\lambda(Z)\to H(Z)$, making Tsallis a
  one-parameter generalization of Shannon entropy.

@@

\textbf{Homogeneous case.} 
Let $Z\sim\Gamma_{\mathrm{SAR}}(\mu, L)$  with pdf given by \eqref{E:gamma1}. Define the integral:
= \int_0^\infty \left[f_{\Gamma_{\text{SAR}}}(z;\mu, L)\right]^{\lambda} \,\mathrm{d}z.
Substituting the density function and simplifying constants, we get
= \left(\frac{L^L}{\Gamma(L)\,\mu^L}\right)^{\lambda}
  \int_0^\infty
     z^{\lambda(L-1)}
     \exp\left(-\frac{\lambda L}{\mu} z\right)\,\mathrm{d}z.
Using the Gamma integral identity
\int_{0}^{\infty} x^{p-1}e^{-q x}\,\mathrm{d}x
&= \frac{\Gamma(p)}{q^{p}}, \quad p,q>0,
with $p = \lambda(L-1) + 1$ and $q = \frac{\lambda L}{\mu}$. This leads to
= \left(\frac{L^L}{\Gamma(L)\,\mu^L}\right)^{\lambda}
  \cdot
  \frac{\Gamma\big(\lambda(L-1)+1\big)}
       {\left(\frac{\lambda L}{\mu}\right)^{\lambda(L-1)+1}}.
After rearranging powers of $L$, $\mu$ and $\lambda$, we obtain
= \frac{L^{\lambda - 1}\,\mu^{1 - \lambda}\;
       \Gamma\big(\lambda(L - 1) + 1\big)}
      {\lambda^{\lambda(L - 1) + 1}\,
       \left[\Gamma(L)\right]^{\lambda}}.
\label{eq:Jgamma1}
Substituting Equation \eqref{eq:Jgamma1} into the Tsallis entropy definition \eqref{eq:tsallis}, we have:
T_\lambda\big(\Gamma_{\text{SAR}}(\mu, L)\big)
= \frac{1}{\lambda - 1}
   \left[
     1 -
     \frac{L^{\lambda - 1}\,\mu^{1 - \lambda}\;
           \Gamma\big(\lambda(L - 1) + 1\big)}
          {\lambda^{\lambda(L - 1) + 1}\;
           \left[\Gamma(L)\right]^{\lambda}}
   \right].
\label{eq:TsallisGammaFinal1}
This expression can also be written in an equivalent form:
<!-- \begin{equation*} -->
<!-- T_\lambda\big(\Gamma_{\text{SAR}}(\mu, L)\big) -->
<!-- = \frac{1}{\lambda - 1} \left[ -->
<!--    1 - -->
<!--    \mu^{1 - \lambda}\, -->
<!--    L^{\lambda - 1}\, -->
<!--    \lambda^{ - [\lambda(L - 1) + 1] }\, -->
<!--    \Gamma\big(\lambda(L - 1) + 1\big)\, -->
<!--    \Gamma(L)^{ - \lambda} -->
<!-- \right], -->
<!-- \end{equation*} -->
\begingroup\small
\label{eq:gammasar-Tsallis}
T_\lambda\bigl(\Gamma_{\mathrm{SAR}}(\mu, L)\bigr)=
\frac{1}{\lambda-1}\Bigl\{1-
+\ln\Gamma\bigl(\lambda(L-1)+1\bigr) \\
-(\lambda(L-1)+1)\ln\lambda
\endgroup

@@

To express $I$ in terms of the mean $\mu$, we use the relation $\gamma = -\mu(\alpha + 1)$ valid for $\alpha < -1$. Substituting into \eqref{eq:J_beta1} gives

@@

This expression extends the Tsallis entropy of the $\Gamma_{\mathrm{SAR}}(\mu, L)$ model \eqref{eq:gammasar-Tsallis} by including additional terms that depend on the texture parameter $\alpha$. It can be written in an exponential-logarithmic form as:
\begingroup\small

@@

  T_\lambda\bigl(\mathcal{G}^0_I(\mu,\alpha,L)\bigr) = T_\lambda\bigl(\Gamma_{\mathrm{SAR}}(\mu, L)\bigr)\;+\;\frac{1}{\lambda-1}\,

@@

\lim_{\alpha \to -\infty} T_\lambda\bigl(\mathcal{G}^0_I( \alpha, \mu, L)\bigr) = T_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr).

@@

Appendix A provides proof of this limiting behavior.

Figure&nbsp;\ref{fig:entropy-diagram} provides a conceptual overview of the decomposition structure of the differential entropy for the heterogeneous model $\mathcal{G}^0_I$. This entropy is expressed as the sum of a baseline term corresponding to the homogeneous $\Gamma_{\text{SAR}}$ distribution and an excess term $\Delta_\alpha$ that reflects the contribution of texture. The decomposition applies to the three entropy measures considered in this study: Shannon \eqref{eq:GIO-Sh1}, Rényi \eqref{eq-HGI0-R}, and Tsallis \eqref{eq:GI0-Tsallis}. In all cases, the excess term depends on the roughness parameter $\alpha$ and vanishes as $\alpha\to-\infty$, thus recovering the homogeneous case.

```{=latex}
\begin{figure}[H]
  \centering
  \caption{Conceptual diagram of entropy decomposition.}
  \label{fig:entropy-diagram}

  \begin{tikzpicture}[
    scale=0.88, every node/.style={transform shape},
    node distance=1.2cm,
    entropy/.style={
      rectangle, rounded corners, draw=blue!60, fill=blue!10, thick,
      minimum width=2.6cm, align=center, font=\sffamily
    },
    property/.style={
      rectangle, draw=orange!60, fill=orange!10, thick,
      minimum width=2.6cm, align=center, font=\sffamily
    },
    arrow/.style={-Stealth, thick}
  ]
